Mozilla Data Science Project - 2019

### 2019-12-06 poc/EDA/EDA.Rmd
updating the R markdown

---
title: "Mozilla Data Science Project - 2019"
author: "Mariana de Oliveira Santos Silva"
date: "December 05, 2019"
output: 
  html_document:
    toc: true
    theme: united
    highlight: textmate
---

@@

## Setting-up

```{r message=FALSE, warning=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
```

@@

To get introduced to our dataset:

```{r}
kable(introduce(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

Let's have a look on the basic information of the dataset.

```{r, fig.width = 10, fig.height = 5}
plot_intro(df_train_f, ggtheme = theme_minimal())
```

Sanity check on the structure and show sample data for each variable.

```{r}
str(df_train_f)
```

When we get a new dataset to analyze, is important to know if there are missing values and the data type. The `df_status` function can help us by showing these numbers in relative and percentage values.

```{r}
kable(df_status(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

@@

@@

`to-do`

<!-- Some variables are encoded as numbers, but they are codes or categories and the models don't handle them in the same way. -->

@@

Yes. Variables with lots of zeros may not be useful for modeling and, in some cases, they may dramatically bias the model.

@@

`to-do`
<!-- Factor/categorical variables with a high number of different values (~30) tend to do overfitting if the categories have low cardinality. -->

## Step 2 : Analyzing Categorical Variables

```{r}
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

freq(df_train_f$label)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution of all discrete variables
plot_bar(df_train_f)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution for release dataframe
plot_bar(df_release)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution beta dataframe
plot_bar(df_beta)
```

### EDA on Coninuous Variables

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_train_f)
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_release)
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_beta)
```

>>>>

---
title: "Mozilla Data Science Project - 2019"
author: "Mariana de Oliveira Santos Silva"
date: "December 05, 2019"
output: 
  html_document:
    toc: true
    theme: united
    highlight: tango
    df_print: kable
---

@@

## Setting-up
```{r message=FALSE, warning=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggthemes)
```

@@

To get introduced to our dataset, et's have a look on the basic information of the dataset.

```{r echo=FALSE}
kable(introduce(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

```{r echo=FALSE, fig.height=10, fig.width=10}
plot_intro(df_train_f, ggtheme = theme_minimal())
```

### Observations

* Most of the dataset is composed of continuous variables
* No NAs values (were handled in preprocessing)

### Structure

Let's use `glimpse` function to display a vertical preview of the dataset. So we can easily preview data type and sample data.

```{r}
glimpse(df_train_f)
```

If we want to get some metrics about data types, zeros, infinite numbers, and missing values, we can use the `df_status` function.

```{r}
kable(df_status(df_train_f, FALSE)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

@@

#### Observations

@@

None. It seems that this has already been dealt with in preprocessing.

@@

Yes. Variables with lots of zeros may not be useful for modeling and, in some cases, they may dramatically bias the model. For example, the `content_crashes` is 100% equal to zero.

@@

Factor/categorical variables with a high number of different values (~30) tend to do overfitting if the categories have low cardinality. Besides `client_id`, it doesn't seem like there are other high cardinality variables.

## Step 2 : Beta vs Release

```{r, fig.width = 7.5}
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

f <- freq(df_train_f$label)
```

```{r}
summary(df_release)
```

```{r}
summary(df_beta)
```

## Step 3 - Analyzing Discrete Variables

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution release dataframe
plot_bar(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution beta dataframe
plot_bar(df_beta, ggtheme = theme_minimal(base_size = 15))
```

## Step 4 - Analyzing Continuos Variables

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15))
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=daily_max_tabs, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Daily Max Tabs") + xlim(0, 20) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=startup_ms, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Startup (ms)") + xlim(0, 5) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=num_active_days, group=label, fill=label)) +
    geom_histogram(stat = "bin", position = "identity", alpha=0.5, main="Num Active Days") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=active_hours, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Active Hours") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=num_pages, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Num Pages") + xlim(0, 50000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=daily_num_sessions_started, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Daily Num Sessions Started") + xlim(0, 20) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=TIME_TO_DOM_INTERACTIVE_MS, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Time to DOM Interactive (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=FX_PAGE_LOAD_MS_2_PARENT, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="FX Page Load (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=TIME_TO_DOM_CONTENT_LOADED_END_MS, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Time to DOM Content Loaded End (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=TIME_TO_LOAD_EVENT_END_MS, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Time to Load Event End (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=TIME_TO_NON_BLANK_PAINT_MS, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Time to Non Blank Paint (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

```{r message=FALSE, warning=FALSE}
ggplot(data=df_train_f, aes(x=TIME_TO_DOM_COMPLETE_MS, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6, main="Time to DOM Complete (ms)") +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum()
```

### Observations
